### 2024-01-01 thesis/chapters/01-introduction.tex
Fix grammar

\chapter{Introduction}
The industry experiences a shift in traditional asset operational status evaluation and utilization. The rise of Industry 4.0 means greater automation and robotization of the production halls to achieve optimal usage of available resources. The secondary aspect in the enterprises' endeavor, however not less important, is to keep track of the equipment wear and tear. The corrective action be it repair or replacement should be done on time in response to the key indicators.

@@

Vibrations are non-intrusive way to sense and record eventually fatal deficiencies right at the onset. The experts use it to distinguish faulty states and to identify the malfunction's root causes. In critical circumstances such as is the case of the large turbines in the power plants, the precautions leading to regular machinery check-ups are already in place. The monitoring solution has to be sufficiently independent, reliable, and accurate to reach wider acceptance and spread.

The main issue to consider in large-scale machinery monitoring with vibrations is that there are lots of uninformative streams of samples not directly useful for the production line operator. The dashboard must aggregate these flows into trend variables with severity levels categorized based on industrial standards. The majority of signals are viewed once at the maximum. Therefore to store or even transmit them from the edge device in its entirety would be wasteful. The complex overview of the mechanical equipment status is attainable only when agent devices and sensors are cheap enough with a long lifespan powered out of the battery pack. Preferably these devices should also remain physically small to reduce the additional clutter in the factory.

Attempted machine learning and deep learning approaches have the crucial impediment that the construction of every single machine is unique to some extent because of tolerances and variable load during regular operation. The model has to be trained for the target environment to achieve the ideal performance. In addition, the failures are relatively rare events that usually occur several months apart. In these circumstances, it is hard to quickly obtain a large enough sample of fault events. Novelty detection is a technique that can be applied in this case.

@@

Chapter \ref{chapter:design} on solution design starts by posing research questions and establishing goals. Then, the MaFaulDa dataset statistical properties and labeling procedure are described on the original samples as well as on extracted attributes. Data rates with feature selection are derived based on the MaFaulDa. In the section about classification models, we compare the performance for various feature sets in the k-nearest neighbors algorithm. Finally, in preparation for the evaluation part, the sensor device is propresed which is suitable for measurements of designated machines in the industrial environment.

>>>>

\chapter{Introduction}
The industry experiences a shift in traditional asset operational status evaluation and utilization. The rise of Industry 4.0 means greater automation and robotization of the production halls to achieve optimal usage of available resources. The secondary aspect in the enterprises' endeavor, but not less important, is to keep track of the equipment's wear and tear. The corrective action, be it repair or replacement, should be done on time in response to the key indicators.

@@

Vibrations are a non-intrusive way to sense and record eventually fatal deficiencies right at the onset. The experts use it to distinguish faulty states and to identify the malfunction's root causes. In critical circumstances such as is the case of the large turbines in power plants, the precautions leading to regular machinery check-ups are already in place. The monitoring solution has to be sufficiently independent, reliable, and accurate to achieve wider acceptance and spread.

The main issue to consider in large-scale machinery monitoring with vibrations is that there are lots of uninformative streams of samples not directly useful for the production line operator. The dashboard must aggregate these flows into trend variables with severity levels categorized based on industrial standards. The majority of signals are viewed once at the maximum. Therefore, to store or even transmit them from the edge device in its entirety would be wasteful. The complex overview of the mechanical equipment status is attainable only when agent devices and sensors are cheap enough with a long lifespan powered out of the battery pack. Preferably these devices should also remain physically small to reduce the additional clutter in the factory.

Attempted machine learning and deep learning approaches have the crucial impediment that the construction of every single machine is unique to some extent because of tolerances and variable load during regular operation. The model has to be trained for the target environment to achieve the ideal performance. In addition, failures are relatively rare events that usually occur several months apart. In these circumstances, it is hard to quickly obtain a large enough sample of fault events. Novelty detection is a technique that can be applied in this case.

@@

Chapter \ref{chapter:design} on solution design starts by posing research questions and establishing goals. Then, the MaFaulDa dataset statistical properties and labeling procedure are described on the original samples as well as on extracted attributes. Data rates with feature selection are derived based on the MaFaulDa. In the section about classification models, we compare the performance for various feature sets in the k-nearest neighbors algorithm. Finally, in preparation for the evaluation part, the sensor device is proprosed which is suitable for measurement of designated machines in the industrial environment.

### 2024-01-01 thesis/chapters/02-analysis-A-CdM.tex
Fix grammar

The defect needs to be repaired or replaced, preferably without significant production downtime, further damage to the other attached elements, or any endangerment of the responsible personnel. The maintenance strategies are chosen according to the machine's importance as a result of its failure effect evaluation on the system. The guide to set appropriate maintenance procedures as outlined in the IEC~60706-2 standard, and involves reliability-centered maintenance analysis~\cite{el-thalji_predictive_2019}.

\subsection{Maintenance strategies}
There are three different approaches to maintenance across the industry: \textbf{reactive, preventive, and predictive}~\cite{scheffer_practical_2004}. In general, the more sophisticated methods are beneficial in a high-stakes environment. The unexpected machine shutdown can have a negative economic impact on the enterprise, resulting in decreased product quality and demands for spare parts to be ready in the supply inventory at all times. In certain situations, it suffices to utilize a simpler maintenance program, but the predictive maintenance gains attraction in the Industry 4.0 to optimize usage of assets~\cite{cinar_machine_2020}.

\textbf{Reactive maintenance} allows machinery to run until a complete failure. This is the most inappropriate way to maintain the production line, but it is straightforward enough. It requires a large stock of replacement parts on-site and breakage inflicts a ``crisis management mode'' upon the plant \cite{scheffer_practical_2004}. On-demand repairs are justified when short downtime is acceptable, full and swift replacement of a broken machine with a backup is possible, or there is a negligible threat to the surrounding environment on failure~\cite{ziaran_technicka_2013}.

\textbf{Preventive maintenance} is performed before any issue is detected. Maintenance occurs at regular intervals derived from a predetermined period in the calendar or expected machine running time (MTTF - Mean Time To Failure). The schedule is crucial but can result in components being replaced in good condition creating waste. The parts can occasionally stay in operation too long, and the machine breaks as a result. Conservative planning is usually the norm to keep the machine always in a perfect state, and therefore more frequent interventions are required~\cite{mohanty_machinery_2015}.

\textbf{Predictive maintenance} known as condition-based maintenance (CbM), improves the predictability of reactive maintenance and eliminates the waste in overall resource utilization of cautious prevention. The machine downtime is scheduled after the detection of unhealthy trends in fault monitoring with sensors and the identification of troublesome components.

A measurable decrease in effectivity allows us to order necessary parts in advance and organize repairs of several machines at a convenient time. The misdetection leads to increased costs compared to previous methods and raises the expectation that faults are distinguishable among themselves~\cite{davies_handbook_2012}.

@@

The \emph{P-F curve} is a widespread representation of equipment degradation over time based on historical records (Fig.~\ref{fig:p-f-curve}). Corrective action should be taken between the event of potential failure (P), when the fault detection is activated, and functional failure point (F) in the P-F interval~\cite{bousdekis_enterprise_2021}.  These division points are not exactly set but have statistical distribution to them.

@@

A high failure rate is present not only at the worn-out stage when the parts are fatigued or corroded but also in the early stages soon after assembly. Manufacturing or material defects, inadequate installation, or improper start-up procedures, are all suspected causes. During the stable middle phase, malfunction can occur after the machine's excessive overload. The time plot to failure rate is known as the bathtub curve~(Fig.~\ref{fig:bathtub-curve}).

\subsection{Vibration fault types}
Mechanical problems during machinery operation bring about vibrations in many instances. Therefore vibroacoustic diagnostics is considered as one of the most important methods in early component fault identification~\cite{ziaran_technicka_2013}.

@@

One of the methods vibration experts utilize in the identification of the damaged part from the frequency spectrum is \textbf{order analysis}. The excessive peaks at harmonic frequencies are of interest. Harmonics are integer multiples of fundamental frequency (1x rpm)~(Tab.~\ref{tab:vibration-causes}):

@@

The vibration severity is the maximum of magnitudes measured in two radial directions (horizontal, and vertical) or supplemented with a third direction along the shaft in the axial axis. Multiple measurement locations, i.e. on different bearings or couplings, should be assessed independently.

Criteria introduced to judge vibration severity are absolute vibration magnitude, change in the magnitude vector, and rate of change. In terms of maximal magnitudes, the machines of varied sizes are split into four severity zones defined in the chart ~(Tab.~\ref{tab:iso20816-vibration-severity}). The values in this table serve as guidelines towards realistic requirements between machine manufacturers and their customers.

@@

\emph{Zone A} is reserved for newly commissioned machines. \emph{Zone B} signifies suitability for long-term operation. In \emph{zone C} the machine is deemed in unsatisfactory condition and corrective action should be taken soon. Finally, in \emph{zone D} vibrations can cause damage to the machine. The span of acceptable values differs with the machine class from I through to IV and their output power of 15~kW (class~I), 75~kW (class~II), 10~MW (class~III), or greater.

The operational limits in the form of \emph{alarms} and \emph{trips} are usually established on the zone boundaries or close to them. Alarms are placed between zones B and C providing a warning about reaching the threshold significant for noticeable change. Trips in between zones C and D urge immediate action or machine shut down. Both limits should not exceed 1.25 times the upper boundary of lower zones and initially are set based on previous experience with the machine~\cite{noauthor_iso_2002}.

\textbf{ISO 13373-1:2002} delves into further nuances of vibration monitoring and expands on procedures outlined in termos of the vocabulary in ISO 20186. According to the standard, the data collection operates in continuous or periodic observation modes which follow an event or interval. Both designs can be permanently mounted. In continuous collection it is recommended to have ``multiplexing rate sufficiently rapid so there is no significant data or trends lost''~\cite{noauthor_iso_2002}. When channels are scanned successively with gaps between data points the system is known as ``scanning''.

The condition monitoring program is run according to a flowchart adapted from one designed in the standard specifically to best benefit the plant. Those steps can be summarized as follows~\cite{noauthor_iso_2002}:
\begin{enumerate}
\itemsep0pt
\item Review machinery history and establish failure modes.
\item When vibration monitoring is not applicable check for other condition-monitoring techniques or resort to preventive maintenance.
\item Select monitoring points and take preliminary vibration measurements.
\item Select vibration monitoring techniques: broadband, frequency analysis, or special techniques, and set parameters of measurement units.
\item Take baseline measurements.
\item Change levels that would warrant investigation.
\item Carry out routine condition monitoring.
\item If an alarm was exceeded, notify appropriate personnel to review data and trends, perform diagnostic evaluation, and repair as necessary. In case a new baseline is needed, continue in the step of taking baseline measurements.
\item Shut down the machine when the trip level is exceeded. Then proceed the same as after the alarm trigger.
\end{enumerate}

Measurement of vibrations should be accompanied by a description of the machine and its operating conditions. The machine description includes its identifier and type, power source, rated rotation speed and power, configuration (shaft or belt driven), and machine support. Measurement parameters are to be recorded alongside the measurement value itself such as timestamp, transducer type, sensor location and orientation in \textbf{MIMOSA convention} (Machinery Information Mangement Open System Alliance), measurement units and units qualifier (p-p, RMS), and other processing options (filters, number of averages, etc.)~\cite{noauthor_iso_2002}.

The transducer of choice for condition monitoring is the accelerometer which can provide the acceleration value of the body and velocity after signal integration. However, standard advise against double integrating for displacement. The recommended frequency range for an accelerometer is 0.1~Hz to 30~kHz. The choice of transducer mount significantly lowers its resonance frequency which is least influenced by stud mount and stiff cement mount. The resonance is reduced to around 8~kHz with the use of soft epoxy or permanent magnet.

@@

Broadband measurement requires ``frequency ranges of 0.2 times the lowest rotational frequency to the highest frequency of interest''~\cite{noauthor_iso_2002}, not exceeding 10 kHz, with RMS velocity 0.1 - 100~mm/s. Bearings and gears diagnosis may push the upper-frequency limit even higher. The tolerances of amplitude and frequency calibrations fall into two types with allowable tolerances of $\pm 5 \%$ or $\pm 10 \%$~(Fig.~\ref{fig:tranducer-response}).

@@

The vibration baseline is defined by broad-band magnitudes and phases of motion vectors, the waveform in the time and frequency domain, the rotational speed of the machine as well as its frequency response to different speeds during start-up and coast-down captured in the Bode plot and waterfall plot. Changes during the machine operation are then depicted in value trends. Trends can be shown of overall amplitudes or limited to frequency bands.

>>>>

The defect needs to be repaired or replaced, preferably without significant production downtime, further damage to the other attached elements, or any endangerment of the responsible personnel. The maintenance strategies are chosen according to the machine's importance as a result of potential failure's impact on the system. The guide to set appropriate maintenance procedures as outlined in the IEC~60706-2 standard, and involves reliability-centered maintenance analysis~\cite{el-thalji_predictive_2019}.

\subsection{Maintenance strategies}
There are three different approaches to maintenance across the industry: \textbf{reactive, preventive, and predictive}~\cite{scheffer_practical_2004}. In general, the more sophisticated methods are beneficial in a high-stakes environment. The unexpected machine shutdown can have a negative economic impact on the enterprise, resulting in decreased product quality and demands for spare parts to be ready in the supply inventory at all times. In certain situations, it suffices to utilize a simpler maintenance program, but predictive maintenance gains attraction in Industry 4.0 to optimize usage of assets~\cite{cinar_machine_2020}.

\textbf{Reactive maintenance} allows machinery to run until a complete failure. This is the most inappropriate way to maintain the production line, but it is straightforward enough. It requires a large stock of replacement parts on-site and breakage inflicts a ``crisis management mode'' upon the plant \cite{scheffer_practical_2004}. On-demand repairs are justified when short downtime is acceptable, full and swift replacement of a broken machine with a backup is possible, or there is a negligible threat to the surrounding environment from failure~\cite{ziaran_technicka_2013}.

\textbf{Preventive maintenance} is performed before any issue is detected. Maintenance occurs at regular intervals derived from a predetermined period in the calendar or expected machine running time (MTTF - Mean Time To Failure). The schedule is crucial but can result in components being replaced in good condition, creating waste. The parts can occasionally stay in operation too long, and the machine breaks as a result. Conservative planning is usually the norm to keep the machines always in a perfect state, and therefore more frequent interventions are required~\cite{mohanty_machinery_2015}.

\textbf{Predictive maintenance} known as condition-based maintenance (CbM), improves the predictability of reactive maintenance and eliminates waste in overall resource utilization of cautious prevention. The machine downtime is scheduled after the detection of unhealthy trends in fault monitoring with sensors and the identification of troublesome components.

A measurable decrease in effectiveness allows us to order necessary parts in advance and organize repairs of several machines at a convenient time. The missed detection leads to increased costs compared to previous methods and raises the expectation that faults are distinguishable among themselves~\cite{davies_handbook_2012}.

@@

The \emph{P-F curve} is a widespread representation of equipment degradation over time based on historical records (Fig.~\ref{fig:p-f-curve}). Corrective action should be taken between the event of potential failure (P), when the fault detection is activated, and functional failure point (F) in the P-F interval~\cite{bousdekis_enterprise_2021}.  These division points are not exactly set but have a statistical distribution to them.

@@

A high failure rate is present not only at the worn-out stage when the parts are fatigued or corroded, but also in the early stages soon after assembly. Manufacturing or material defects, inadequate installation, or improper start-up procedures, are all suspected causes. During the stable middle phase, malfunction can occur after the machine's excessive overload. The time plot to failure rate is known as the bathtub curve~(Fig.~\ref{fig:bathtub-curve}).

\subsection{Vibration fault types}
Mechanical problems during machinery operation bring about vibrations in many instances. Therefore, vibroacoustic diagnostics is considered as one of the most important methods in early component fault identification~\cite{ziaran_technicka_2013}.

@@

One of the methods vibration experts utilize in the identification of the damaged part, according to the frequency spectrum, is \textbf{order analysis}. The excessive peaks at harmonic frequencies are of interest. Harmonics are integer multiples of fundamental frequency (1x rpm)~(Tab.~\ref{tab:vibration-causes}):

@@

The vibration severity is the maximum magnitude measured in two radial directions (horizontal, and vertical) or supplemented with a third direction along the shaft on the axial axis. Multiple measurement locations, i.e. on different bearings or couplings, should be assessed independently.

Criteria introduced to judge vibration severity are absolute vibration magnitude, change in the magnitude vector, and rate of change. In terms of maximal magnitudes, the machines of varying sizes are split into four severity zones defined in the chart ~(Tab.~\ref{tab:iso20816-vibration-severity}). The values in this table serve as guidelines towards realistic requirements between machine manufacturers and their customers.

@@

Zone A is reserved for newly commissioned machines. Zone B signifies suitability for long-term operation. In zone C the machine is deemed in unsatisfactory condition and corrective action should be taken soon. Finally, in zone D vibrations can cause damage to the machine. The span of acceptable values differs with the machine class from \rom{1} through to \rom{4} and their output power of 15~kW (class~\rom{1}), 75~kW (class~\rom{2}), 10~MW (class~\rom{3}), or greater.

The operational limits in the form of \emph{alarms} and \emph{trips} are usually established on the zone boundaries or close to them. Alarms are placed between zones B and C providing a warning about reaching the threshold significant for noticeable change. Trips in between zones C and D urge immediate action or machine shut down. Both limits should not exceed 1.25 times the upper boundary of the lower zones and initially are set based on previous experience with the machine~\cite{noauthor_iso_2002}.

\textbf{ISO 13373-1:2002} delves into further nuances of vibration monitoring and expands on procedures outlined in terms of the vocabulary in ISO 20186. According to the standard, the data collection operates in continuous or periodic observation modes which follow an event or interval. Both designs can be permanently mounted. In continuous collection, it is recommended to have a ``multiplexing rate sufficiently rapid, so there is no significant data or trends lost''~\cite{noauthor_iso_2002}. When channels are scanned in succession with gaps between data points, the system is known as ``scanning''.

The condition monitoring program is run according to a flowchart adapted from one designed by the standard specifically to best benefit the plant. Those steps can be summarized as follows~\cite{noauthor_iso_2002}:

\begin{enumerate}
\itemsep0pt
\item Review machinery history and establish failure modes.
\item When vibration monitoring is not applicable, check for other condition-monitoring techniques or resort to preventive maintenance.
\item Select monitoring points and take preliminary vibration measurements.
\item Select vibration monitoring techniques: broadband, frequency analysis, or special techniques, and set parameters of measurement units.
\item Take baseline measurements.
\item Change levels that would warrant investigation.
\item Carry out routine condition monitoring.
\item If an alarm was exceeded, notify appropriate personnel to review data and trends, perform diagnostic evaluation, and repair as necessary. In case a new baseline is needed, continue in the step of taking baseline measurements.
\item Shut down the machine when the trip level is exceeded. Then proceed the same as after the alarm is triggered.
\end{enumerate}

Measurement of vibrations should be accompanied by a description of the machine and its operating conditions. The machine description includes its identifier and type, power source, rated rotation speed and power, configuration (shaft or belt driven), and machine support. Measurement parameters are to be recorded alongside the measurement value itself, such as timestamp, transducer type, sensor location and orientation in \textbf{MIMOSA convention} (Machinery Information Management Open System Alliance), measurement units and units qualifier (p-p, rms), and other processing options (filters, number of averages, etc.)~\cite{noauthor_iso_2002}.

The transducer of choice for condition monitoring is the accelerometer, which can provide the acceleration value of the body and velocity after signal integration. However, standard advise against double integrating for displacement. The recommended frequency range for an accelerometer is 0.1~Hz to 30~kHz. The choice of transducer mount significantly lowers its resonance frequency, which is least influenced by the stud mount and stiff cement mount. The resonance is reduced to around 8~kHz with the use of soft epoxy or permanent magnet.

@@

Broadband measurement requires ``frequency ranges of 0.2 times the lowest rotational frequency to the highest frequency of interest''~\cite{noauthor_iso_2002}, not exceeding 10 kHz, with RMS velocity 0.1 - 100~mm/s. Bearings and gear diagnosis may push the upper-frequency limit even higher. The tolerances of amplitude and frequency calibrations fall into two types with allowable tolerances of $\pm 5 \%$ or $\pm 10 \%$~(Fig.~\ref{fig:tranducer-response}).

@@

The vibration baseline is defined by broad-band magnitudes and phases of motion vectors, the waveform in the time and frequency domain, the rotational speed of the machine as well as its frequency response to different speeds during start-up and coast-down captured in the Bode plot and waterfall plot. Changes during the machine's operation are then depicted in value trends. Trends can be shown by overall amplitudes or limited to frequency bands.
